---
title: "Assignment 3 - Causal inference"
author: "RF"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Exploring causal inference issues

In this assignment we explore some issues related to multiple regressions (regressions with more than one predictor), and inferred (causal) relations between variables. N.B. the data is simulated (to make sure I know the actual mechanism generating it), but it's based on a real study. So bear with a longish introduction to get into the details of what we are doing and why it is important.

### Altercentric intrusion in schizophrenia

People with schizophrenia often report altered control and distinction of self-other representations: intrusive thoughts, hearing of voices, delusions of mind reading, paranoia, etc (a substantial portion of the psychotic symptoms experienced in schizophrenia). These have been variously attributed to hypermentalizing (over attribution of mental states to others), social impairment (over preoccupation with own thought processes), hyper socialization (inability to inhibit information from others), etc.

The current study investigates 1) whether schizophrenia is indeed related to altered control and distinction of self-other representations, in particular altercentric intrusions (inability to inhibit social information), and 2) whether these are related to the relevant psychotic symptoms. N.B. the actual study also investigates egocentric intrusion, do check the papers below if interested.

The task is a slightly modified version of this: https://www.ncbi.nlm.nih.gov/pubmed/20731512 You look at a picture with some dots visible to you, as well as with a different person with a different set of dots visible to them. The number of dots you see and that the other sees can be the same (congruent condition) or not (incongruent condition). You are tasked to indicate whether a given number (e.g. 3) matches the number of dots you see (and the dots visible to the other person are irrelevant to the task).


The tasks investigates altercentric intrusion: will your reaction time change according to whether the other person is seeing the same amount of dots as you, or not? The idea is that if you correctly inhibit social information, your reaction time should not change, as the information about the other person is not relevant. On the contrary, if you nevertheless use task irrelevant social information, you'll be slower at indicating whether 3 is the right number of dots when the other person sees a different amount of dots than you (conflicting information).
The bigger the difference between RTs in the congruent and incongruent condition the bigger the altercentric intrusion effect.

For each participant you have 6 variables: 1) ID, 2) AltercentricIntrusion (continuous score), 3) Diagnosis (schizophrenia vs. control), 4) VoiceHearing (severity of voice hearing symptoms, continuous score of the severity of the symptom as measured by a clinician), 5) MindReading (severity of delusions of mind reading, continuous score of the severity of the symptom as measured by a clinician); 6) Apathy (severity of lack of motivation in taking care of oneself, from washing to showing up at work, continuous score of the severity of the symptom as measured by a clinician).

The research questions you have to answer are the following:

## First part

Q1.1) Does schizophrenia involve altercentric intrusion? Define model and priors. Test the implications of your priors (prior predictive checks) and if needed adjust them. Run the model. Test the quality of the fitted model (posterior predictive checks). Assess the evidence in favor of an increased altercentric intrusion in schizophrenia. Report the model and the results, including plots.

```{r}
# Loading packages
library(pacman)
pacman::p_load(tidyverse, brms)


# Preparing the data
d <- read_csv("Ass3.csv")

summary(d)

# Diagnosis in words (control/schizophrenia) instead of factor (0/1)
d$Diagnosis <- plyr::revalue(as.character(d$Diagnosis), 
                             c("0"="Controls", "1"="Schizophrenia"))

# Making ID and Diagnosis factors
d <- d %>%
  mutate(
    ID = as.factor(ID),
    Diagnosis = as.factor(Diagnosis)
  )


# Defining the formula/model (bf = bayesian formula)
AltercentricDiagnosis_f1 <- bf(AltercentricIntrusion ~ 1 + Diagnosis) # Gives one intercept and one estimate

AltercentricDiagnosis_f <- bf(AltercentricIntrusion ~ 0 + Diagnosis) # Gives the estimate for both controls and schizophrenia (DEN VI BRUGER)


# Designing the priors
# Telling us which priors we have to give
get_prior(AltercentricDiagnosis_f1, family = gaussian, d)
get_prior(AltercentricDiagnosis_f, family = gaussian, d)

# Looking at the data
summary(d$AltercentricIntrusion)

# Defining priors
priorDiagnosis <- c(
  prior(normal(4, 1), class = b), # First prior is for the beta with a normal distribution - 4 and 1 are based on summary of alt_in (mean = 4 in data, max = 6, so SD = 1, since 2 SD makes it go from 4 to 6)
  prior(normal(1, 2), class = sigma) # SD of alt_in is 1, so we expect that kind of error with SD of 2, it can error up to 5
) 


# Testing the priors (doublecheck whether they make any sense)
AltercentricDiagnosis_PriorCheck_m <- brm(
  formula = AltercentricDiagnosis_f,
  data = d,
  family = gaussian,
  prior = priorDiagnosis,
  sample_prior = "only" # We sample the prior because we want to test the prior
)

# Prior predictive check
# Given only the prior and not the data - are going to look like (with 100 simulations of the model)
pp_check(AltercentricDiagnosis_PriorCheck_m, nsamples = 100)


## Fitting the actual model
AltercentricDiagnosis_m <- brm(
  formula = AltercentricDiagnosis_f,
  data = d,
  family = gaussian,
  prior = priorDiagnosis,
  sample_prior = T # To set up the actual model
)


# Posterior predictive check (plot)
pp_check(AltercentricDiagnosis_m, nsamples = 100)


## Checking the model for warnings (and getting estimates)
AltercentricDiagnosis_m


# Hypothesis testing + updating check
# To see whether the model has learned anything - comparing the prior and the posterior
plot(hypothesis(AltercentricDiagnosis_m,
           "DiagnosisSchizophrenia > DiagnosisControls"))
  # Prior for the difference between controls and scz (uncertain). We gave the same prior to both conditions, so it is centred around 0 and quite spread (but it covers the regions we want it to).
  # Posterior is much more certain. It has learned from the data to be much more confident
  # RF is satisfied at this point.

# Testing specific hypothesis (getting estimates)
hypothesis(AltercentricDiagnosis_m,
           "DiagnosisSchizophrenia > DiagnosisControls")
  # Hypothesis that SCZ has a higher alt_in than controls says that there is a difference of 0.36, estimated error (SD) of 0.12 (much smaller than the estimate = good), CL above 0 = good, evidence ratio = high (how many samples are incompatible with our hypothesis = a count --> a ratio. We have 4000???). Overwhelming evidence given the data and the model that scz has a higher level of alt_in.


# Plotting the results (group level)
conditional_effects(AltercentricDiagnosis_m)

# Plotting the results (individual leve)
plot(conditional_effects(AltercentricDiagnosis_m), points = T)
# By looking at a single individual, you would not be able to infer/predict much about diagnose just from alt_in, but at a population level we see a clear effect.

```

The model indicates a credible difference in altercentric intrusion in the two groups supporting our hypothesis (b = 0.36, CIs = 0.17, 0.56, ER = 1332/570.43). Controls showed on average an altercentric intrusion effect of 3.86 (CIs 3.75, 3.98), and schizophrenia of 4.22 (CIs = 4.02, 4.44).
[Add plot of the effects]

# SI
The model had no divergences, a Rhat of 1, and Effective Sample Sizes above 2000 for both Bulk and Tail.
[Add prior and posterior checks plots; add updating check plot]

Q1.2) Is altercentric intrusion related to specific symptoms *in the patients*? Identify which of the symptoms could be relevant. Should you include more than one symptom? Build models, priors, predictive checks. Assess the evidence and report models and results, including plots. Discuss whether the results make sense.
```{r}
# Subsetting the patients only (diagnosed with schizophrenia)
scz <- d %>% 
  filter(Diagnosis == "Schizophrenia")


# Defining formulas/models
f_voice <- bf(AltercentricIntrusion ~ 1 + VoiceHearing)

f_mind <- bf(AltercentricIntrusion ~ 1 + MindReading)

f_apathy <- bf(AltercentricIntrusion ~ 1 + Apathy)

f_voice_mind <- bf(AltercentricIntrusion ~ 1 + VoiceHearing + MindReading)

# f_voice_mind_apathy <- bf(AltercentricIntrusion ~ 1 + VoiceHearing + MindReading + Apathy)


# Designing the priors
# Getting info
get_prior(f_voice, family = gaussian, scz)
summary(scz$VoiceHearing)

get_prior(f_mind, family = gaussian, scz)
summary(scz$MindReading)

get_prior(f_apathy, family = gaussian, scz)
summary(scz$Apathy)

get_prior(f_voice_mind, family = gaussian, scz)

# get_prior(f_voice_mind_apathy, family = gaussian, scz)


# Defining the priors
prior_voice <- c(
  prior(normal(0, 0.5), class = b), # We expect VoiceHearing to positively affect AltercentricIntrusion
  prior(normal(4, 1), class = Intercept),
  prior(normal(1, 2), class = sigma)
)

prior_mind <- c(
  prior(normal(0, 0.5), class = b), # We expect MindReading to positively affect AltercentricIntrusion
  prior(normal(4, 1), class = Intercept),
  prior(normal(1, 2), class = sigma)
)

prior_apathy <- c(
  prior(normal(0, 0.05), class = b), # We do not expect apathy to have any effect, thus being certain that a slope is close to 0
  prior(normal(4, 1), class = Intercept),
  prior(normal(1, 2), class = sigma)
)

prior_voice_mind <- c(
  prior(normal(0, 0.5), class = b), # We expect VoiceHearing and MindReading to positively affect AltercentricIntrusion
  prior(normal(4, 1), class = Intercept),
  prior(normal(1, 2), class = sigma)
)

# prior_voice_mind_apathy <- c(
#   prior(normal(0, 0.5), class = b),
#   prior(normal(4, 1), class = Intercept),
#   prior(normal(1, 2), class = sigma)
# )


# Testing the priors
# Voice
voice_prior_check <- brm(
  formula = f_voice,
  data = scz,
  family = gaussian,
  prior = prior_voice,
  sample_prior = "only"
)

# Prior predictive check
pp_check(voice_prior_check, nsamples = 100)

# Mind
mind_prior_check <- brm(
  formula = f_mind,
  data = scz,
  family = gaussian,
  prior = prior_mind,
  sample_prior = "only"
)

# Prior predictive check
pp_check(mind_prior_check, nsamples = 100)

# Apathy
apathy_prior_check <- brm(
  formula = f_apathy,
  data = scz,
  family = gaussian,
  prior = prior_apathy,
  sample_prior = "only"
)

# Prior predictive check
pp_check(apathy_prior_check, nsamples = 100)

# Voice, mind
voice_mind_prior_check <- brm(
  formula = f_voice_mind,
  data = scz,
  family = gaussian,
  prior = prior_voice_mind,
  sample_prior = "only"
)

# Prior predictive check
pp_check(voice_mind_prior_check, nsamples = 100)

# # Voice, mind, apathy
# voice_mind_apathy_prior_check <- brm(
#   formula = f_voice_mind_apathy,
#   data = scz,
#   family = gaussian,
#   prior = prior_voice_mind_apathy,
#   sample_prior = "only"
# )
# 
# # Prior predictive check
# pp_check(voice_mind_prior_check, nsamples = 100)


# Fitting the model
# Voice
m_voice <- brm(
  formula = f_voice,
  data = scz,
  family = gaussian,
  prior = prior_voice,
  sample_prior = T
)

# Posterior predictive check
pp_check(m_voice, nsamples = 100)

# Mind
m_mind <- brm(
  formula = f_mind,
  data = scz,
  family = gaussian,
  prior = prior_mind,
  sample_prior = T
)

# Posterior predictive check
pp_check(m_mind, nsamples = 100)

# Apathy
m_apathy <- brm(
  formula = f_apathy,
  data = scz,
  family = gaussian,
  prior = prior_apathy,
  sample_prior = T
)

# Posterior predictive check
pp_check(m_apathy, nsamples = 100)

# Voice, mind
m_voice_mind <- brm(
  formula = f_voice_mind,
  data = scz,
  family = gaussian,
  prior = prior_voice_mind,
  sample_prior = T
)

# Posterior predictive check
pp_check(m_voice_mind, nsamples = 100)

# # Voice, mind, apathy
# m_voice_mind_apathy <- brm(
#   formula = f_voice_mind_apathy,
#   data = scz,
#   family = gaussian,
#   prior = prior_voice_mind_apathy,
#   sample_prior = T
# )
# 
# # Posterior predictive check
# pp_check(m_voice_mind_apathy, nsamples = 100)


# Getting estimates and checking the model for warnings
m_voice
#              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept        4.06      0.31     3.47     4.66 1.00     3723     2824
# VoiceHearing     0.07      0.13    -0.18     0.34 1.00     3724     2861

m_mind
#             Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept       4.09      0.21     3.67     4.50 1.00     4090     2876
# MindReading     0.08      0.10    -0.13     0.28 1.00     3777     2921

m_apathy
#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept     4.28      0.14     4.01     4.55 1.00     3772     2960
# Apathy       -0.03      0.05    -0.12     0.06 1.00     3901     2616

m_voice_mind
#              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept        3.75      0.45     2.89     4.62 1.00     3506     2612
# VoiceHearing     0.12      0.14    -0.16     0.40 1.00     3863     3069
# MindReading      0.11      0.12    -0.12     0.34 1.00     3616     2920


# Hypothesis testing + updating check
# Voice
plot(hypothesis(m_voice,
           "VoiceHearing > 0"))

hypothesis(m_voice,
           "VoiceHearing > 0")
# Estimate    0.07
# Est.Error   0.13
# CI.Lower    -0.14     90%-CI for one-sided hypotheses
# CI.Upper    0.29      90%-CI for one-sided hypotheses
# Evid.Ratio  2.42
# post_prob   0.71

# Mind
plot(hypothesis(m_mind,
           "MindReading > 0"))

hypothesis(m_mind,
           "MindReading > 0")
# Estimate    0.08
# Est.Error   0.1
# CI.Lower    -0.09     90%-CI for one-sided hypotheses
# CI.Upper    0.24      90%-CI for one-sided hypotheses
# Evid.Ratio  3.44
# post_prob   0.78

# Apathy
plot(hypothesis(m_apathy,
           "Apathy = 0")) # =

hypothesis(m_apathy,
           "Apathy = 0") # =
# Estimate    -0.03
# Est.Error   0.05
# CI.Lower    -0.12     95%-CI for two-sided hypotheses
# CI.Upper    0.06      95%-CI for two-sided hypotheses
# Evid.Ratio  0.91
# post_prob   0.48

# Voice, mind
plot(hypothesis(m_voice_mind,
           "VoiceHearing > 0"))
plot(hypothesis(m_voice_mind,
           "MindReading > 0"))

hypothesis(m_voice_mind,
           "VoiceHearing > 0")
# Estimate    0.12
# Est.Error   0.14
# CI.Lower    -0.11     90%-CI for one-sided hypotheses
# CI.Upper    0.35      90%-CI for one-sided hypotheses
# Evid.Ratio  4.15
# post_prob   0.81
hypothesis(m_voice_mind,
           "MindReading > 0")
# Estimate    0.11	0.12	-0.08	0.3	5.33	0.84
# Est.Error   0.12
# CI.Lower    -0.08     90%-CI for one-sided hypotheses
# CI.Upper    0.3       90%-CI for one-sided hypotheses
# Evid.Ratio  5.33
# post_prob   0.84


# Plotting the results
# Group level
conditional_effects(m_voice)
conditional_effects(m_mind)
conditional_effects(m_apathy)
conditional_effects(m_voice_mind)

# Individual level
plot(conditional_effects(m_voice), points = T)
plot(conditional_effects(m_mind), points = T)
plot(conditional_effects(m_apathy), points = T)
plot(conditional_effects(m_voice_mind), points = T)
```


## Second part

Q2.1) However, we know that the diagnosis is based on symptom assessment: if the overall sum of symptoms is severe enough, the participant gets a diagnosis. In other words, by selecting the patients, and including the symptoms in the model we might have inadvertently introduced an issue in our inference. Do try to draw a causal graph (Directed Acyclical Graph) of the variables and compare it with the types of causal graphs presented in the slides. Discuss which biases you might have introduced.

Q2.2.) Redesign your analysis following the graph and report how the results change


## Third part

These issues are very difficult to think through, and not knowing the causal mechanisms generating the data in advance makes our inferences even more unreliable. To explore these issues, I recommend using simulations. In other words, defining a "true" model, generating data from it and assessing what different analyses would lead you to infer (and therefore which biases they might introduce). You can find the code I used to simulate your data below.

Q3.1) Look through the code and identify whether the results you have match the underlying truth. Discuss what you have learned.

Q3.2) OPTIONAL: is this a general pattern? Try varying the parameters (e.g. correlation values) and assess whether the new dataset(s) leads to the same biases in your analysis.



```{r}
pacman::p_load(MASS, tidyverse, psych)

seed <- 1981 # Defining a seed so the results are always the same
n <- 300 # Defining the amount of participants

SymptomCorr <- .2 # Defining the correlation of symptoms (as they tend to co-occur)
EffectCorrRel <- .2 # Defining the correlation between relevant symptoms and effect (Some symptoms are positively correlated with the effect)
EffectCorrIrrel <- 0 # Defining the correlation between irrelevant symptoms and effect (none)

# Creating the variance-covariance matrix for the variables we want to generate (3 symptoms, 1 effect)
Sigma <- matrix(data=c(1,SymptomCorr,SymptomCorr,EffectCorrRel,
                       SymptomCorr,1,SymptomCorr,EffectCorrRel,
                       SymptomCorr,SymptomCorr,1,EffectCorrIrrel,
                       EffectCorrRel,EffectCorrRel,EffectCorrIrrel,1),
                       nrow=4,ncol=4)

## Generate data from a multivariate (mvr) normal (n) distribution
d <- mvrnorm(n = n, # number of participant
        mu = c(1.2, 1.2, 1.2, 4), # mean of each variable
        Sigma) # variance co-variance matrix

# Giving meaningful names to variables and add ID
d <- data.frame(
  VoiceHearing = d[,1], 
  MindReading =  d[,2],
  Apathy =  d[,3], 
  AltercentricIntrusion = d[,4],
  ID = seq(nrow(d)))

# Assessing whether the participant has schizophrenia (high enough sum of symptoms)
# Here we choose participants scoring above 75% percentile (the most severe ones)
d$Diagnosis <- 0
d$Diagnosis[(d$VoiceHearing + d$MindReading + d$Apathy) > 
              quantile(d$VoiceHearing + d$MindReading + d$Apathy, .75)] <-1

## Plotting the relation between variables in schizophrenia
d1 <- d %>% subset(Diagnosis==1) %>% dplyr::select(-Diagnosis, -ID)
pairs.panels(d1)

## Plotting the relation between variables all participants
pairs.panels(dplyr::select(d,-Diagnosis, -ID))

write_csv(d, "data/Ass3.csv")
```

